---
title: "Prediction Practice"
author: "Mo Azar"
date: "2024-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prediction Practice for Final Exam Assignemnt

in this project, as given by the coursera we should use data of accelometers on the belt,forearm,arm and dumbell of 6 participants. they were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
out target is to predict the manner in which they did the excercise. 
The outcome variable is classe, a factor variable with 5 levels. For this data set, participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in 5 different fashions:

exactly according to the specification (Class A)

throwing the elbows to the front (Class B)

lifting the dumbbell only halfway (Class C)

lowering the dumbbell only halfway (Class D)

throwing the hips to the front (Class E)

## Installing all required Packages
```{r Install Packages}
library(knitr)
library(lattice)
library(ggplot2)
library(caret)
library(rpart)
library(rpart.plot)
library(corrplot)
library(RColorBrewer)
library(rattle)
```
# setting a random seed
```{r set seed}
set.seed(1362)
```

#Loading the data from csv
```{r loading data}
traincsv <- read.csv("C:\\MLProject\\data\\pml-training.csv")
testcsv <- read.csv("C:\\MLProject\\data\\pml-testing.csv")
```
# cleaning the data
```{r cleaning data}
traincsv <- traincsv[,colMeans(is.na(traincsv)) < .9] # removing na
traincsv <- traincsv[,-c(1:7)] # removing unneccesary columns

# removing zero var
nvz <- nearZeroVar(traincsv)
traincsv <- traincsv[,-nvz]
dim(traincsv)
```

# creating validation and training set
```{r validating&test}
inTrain <- createDataPartition(y=traincsv$classe, p=0.7, list=F)
train <- traincsv[inTrain,]
valid <- traincsv[-inTrain,]
```
# executing models
```{r Executing models}
print("Training models, please wait...")
control <- trainControl(method="cv", number=3, verboseIter=F) #set up fixed training parameters
```
# Models
```{r Models}

mod_trees <- train(classe~., data=train, method="rpart", trControl = control, tuneLength = 5)
pred_trees <- predict(mod_trees, valid)
cmtrees <- confusionMatrix(pred_trees, factor(valid$classe))
fancyRpartPlot(mod_trees$finalModel)


mod_rf <- train(classe~., data=train, method="rf", trControl = control, tuneLength = 5)
pred_rf <- predict(mod_rf, valid)
cmrf <- confusionMatrix(pred_rf, factor(valid$classe))


mod_gbm <- train(classe~., data=train, method="gbm", trControl = control, tuneLength = 5, verbose = F)
pred_gbm <- predict(mod_gbm, valid)
cmgbm <- confusionMatrix(pred_gbm, factor(valid$classe))


mod_svm <- train(classe~., data=train, method="svmLinear", trControl = control, tuneLength = 5, verbose = F)
pred_svm <- predict(mod_svm, valid)
cmsvm <- confusionMatrix(pred_svm, factor(valid$classe))


print("Accuracy and out of sample error rate for each model...")
models <- c("Tree", "RF", "GBM", "SVM")
accuracy <- round(c( cmtrees$overall[1], cmrf$overall[1], cmgbm$overall[1], cmsvm$overall[1]),3) #accuracy
oos_error <- 1 - accuracy #out of sample error
data.frame(accuracy = accuracy, oos_error = oos_error, row.names = models)
```

# running the best model 
```{r run the best model }

print("Using best model to predict class outcome for each test set observation...")
pred <- predict(mod_rf, testcsv)
print(pred)

print("Done.")
```
